Differential geometric algebra with Leibniz and Grassmann

### 2019-11-04 paper/paper.tex
created projective unsplitter norm and tidelift enterprise referral

\begin{theorem}[Integration by parts \& Stokes]
	Let $ \nabla \in\Omega_1 V $ be a Leibnizian vector field operator, then $d,-\partial$ are Hilbert adjoint Hodge-DeRahm operators with $\gen{\,\ast\,}$
	\begin{align*}
		\int_M d\omega\wedge\star\eta +\int_M \omega\wedge\star\partial\eta &= 0, & \gen{d\omega\ast\eta} &=\gen{\omega\ast-\partial\eta}.
	\end{align*}
\end{theorem}
\begin{proof}
	$\partial\omega = \omega\cdot\nabla = \star(\star\omega\wedge\star^2\nabla) = (-1)^n(-1)^{nk}\star d\star\omega$.

	Then  substitute this into $\int_M \omega\wedge(-1)^{mk+m+1}\star\star d\star\eta = (-1)^{km+m+1}(-1)^{(m-k+1)(k-1)}\int_M\omega\wedge d\star\eta$,
	with identity $(-1)^{km+m+1}(-1)^{(m-k+1)(k-1)}=(-1)^k$ and
	$ (-1)^k\int_M\omega\wedge d\star\eta = \int_M d(\omega\wedge\star\eta) - (-1)^{k-1}\omega\wedge d\star\eta = \int_M d\omega\wedge\star\eta$.
	Stokes identity can be proved by relying on a variant of the \textit{common factor theorem} by Browne \cite{browne}.
\end{proof}
\begin{theorem}[Clifford-Dirac-Laplacian]
	The Dirac operator \cite{garling} is $ (\nabla^2)^\frac12\omega = \pm\nabla\ominus\omega = \pm\nabla\wedge\omega \pm \nabla\cdot\omega  = \pm d\omega\pm\partial\omega$.
	\begin{align*}
		\nabla^2\omega &= \nabla\wedge(\omega\cdot\nabla) + (\nabla\wedge\omega)\cdot\nabla) = \mp(\mp\omega\ominus\nabla)\ominus\nabla).
	\end{align*}
	Elements $\omega$ are \textit{harmonic} if $\nabla\omega = 0$ and both \textit{exact} $d\omega=0$ and \textit{coexact} $\delta\omega=0$, such that $\mc H^p M = \setc{\nabla\omega = 0}{\omega\in \Omega^pM}$.
	Hodge \cite{ivancevic}:
	$\Omega^pM=\mc H^pM\oplus\im{d\Omega^{p-1}M}\oplus\im{\partial\Omega^{p+1}M}$.
	Note: $\nabla\omega=-\omega\nabla, \nabla^2\omega=\omega\nabla^2$ for higher-order tensor fields!
\end{theorem}

@@

\begin{lstlisting}[language = Julia]
using Grassmann, Makie
basis"2" # Euclidean
streamplot(vectorfield(v1*exp(π*v12/2)),-1.5..1.5,-1.5..1.5)
streamplot(vectorfield(exp((π/2)*v12/2)),-1.5..1.5,-1.5..1.5)
streamplot(vectorfield(exp((π/4)*v12/2)),-1.5..1.5,-1.5..1.5)
streamplot(vectorfield(v1*exp((π/4)*v12/2)),-1.5..1.5,-1.5..1.5)
@basis S"+-" # Hyperbolic
streamplot(vectorfield(exp((π/8)*v12/2)),-1.5..1.5,-1.5..1.5)
streamplot(vectorfield(v1*exp((π/4)*v12/2)),-1.5..1.5,-1.5..1.5)
\end{lstlisting}
\begin{figure}[ht]
	\centering
	%\caption{Euclidean vector fields based on exponential operator:}
	\begin{subfigure}[b]{0.23\textwidth}
		\includegraphics[width=\textwidth]{img/plane-1.png}
		\caption{$x\oslash e^{\pi v_{12}/2}$, $\mb R^2$}
		%\label{fig:sample_figure}
	\end{subfigure}
	~
	\begin{subfigure}[b]{0.23\textwidth}
		\includegraphics[width=\textwidth]{img/plane-2.png}
		\caption{$ x\oslash e^{\frac\pi2 v_{12}/2}$, $\mb R^2$}
		%\label{fig:sample_figure}
	\end{subfigure}

@@

As a result of Grassmann's exterior \& interior products, the Hodge-DeRahm chain complex from cohomology theory is
\begin{align*}
	0 &\os d{\us\partial\rightleftarrows} \Omega^0(M) \os d{\us\partial\rightleftarrows} \Omega^1(M) \os d{\us\partial\leftrightarrows} \cdots \os d{\us\partial\rightleftarrows} \Omega^n(M) \os d{\us\partial\rightleftarrows} 0,
\end{align*}
having dimensional equivalence brought by the Grassmann-Poincare-Hodge complement duality,
\begin{align*}
	\mc H^{n-p}M &\cong \frac{\ker(d\Omega^{n-p}M)}{\im{d\Omega^{n-p-1}M}}, & \dim\mc H^pM &= \dim\frac{\ker(\partial\Omega^pM)}{\im{\partial\Omega^{p+1}M}}
\end{align*}
The rank of the grade $p$ boundary incidence operator is
$$ \text{rank}\gen{\partial\gen{M}_{p+1}}_p = \min\set{\dim\gen{\partial\gen{M}_{p+1}}_p,\dim\gen{M}_{p+1}} $$
Invariant topological information can be computed using the rank of homology groups, where $b_p(M)=\dim\mc H^pM$ are
$$ b_p(M) = \dim\gen{M}_{p+1} - \text{rank}\gen{\partial\gen{M}_{p+1}}_p - \text{rank}\gen{\partial\gen{M}_{p+2}}_{p+1} $$
Betti numbers with Euler characteristic $\chi(M) = \sum_p (-1)^pb_p$.

@@

In Fig. 4, different possible discrete bivector topologies in a projective Riemann sphere setting are examined. 
The figures are based on the product topology of two rotation bivectors.
When the Euclidean $\mb R^4$ basis is combined with projective geometric algebra, resulting one parameter Lie groups can be visualized in the form of a torus in $\mb R^3$. 
When the fourth $v_\infty$ basis direction is rotated into the Minkowski plane, the double rotation becomes a helix with translational single rotation. In examples (b)$\sim$(c) the bivector is modulated.

>>>>

\begin{theorem}[Integration by parts \& Stokes]
	Let $ \nabla \in\Omega_1 V $ be a Leibnizian vector field operator, then $d,-\partial$ are Hilbert adjoint Hodge-DeRahm operators with $\gen{\,\ast\,}$
	\begin{align*}
		\int_M d\omega\wedge\star\eta +\int_M \omega\wedge\star\partial\eta &= 0, & \gen{d\omega\ast\eta} &=\gen{\omega\ast-\partial\eta}.
	\end{align*}
\end{theorem}
\begin{proof}
	$\partial\omega = \omega\cdot\nabla = \star\inv(\star\omega\wedge\star^2\nabla) = (-1)^n(-1)^{nk}\star d\star\omega$.

	Then  substitute this into $\int_M \omega\wedge(-1)^{mk+m+1}\star\star d\star\eta = (-1)^{km+m+1}(-1)^{(m-k+1)(k-1)}\int_M\omega\wedge d\star\eta$,
	apply the identity $(-1)^{km+m+1}(-1)^{(m-k+1)(k-1)}=(-1)^k$ and
	$ (-1)^k\int_M\omega\wedge d\star\eta = \int_M d(\omega\wedge\star\eta) - (-1)^{k-1}\omega\wedge d\star\eta = \int_M d\omega\wedge\star\eta$.
	Stokes identity can be proved by relying on a variant of the \textit{common factor theorem} by Browne \cite{browne}.
\end{proof}
\begin{theorem}[Clifford-Dirac-Laplacian]
	The Dirac operator \cite{garling} is $ (\nabla^2)^\frac12\omega = \pm\nabla\ominus\omega = \pm\nabla\wedge\omega \pm \nabla\cdot\omega  = \pm d\omega\pm\partial\omega$.
	\begin{align*}
		\nabla^2\omega &= \nabla\wedge(\omega\cdot\nabla) + (\nabla\wedge\omega)\cdot\nabla) = \mp(\mp\omega\ominus\nabla)\ominus\nabla).
	\end{align*}
	Elements $\omega$ are \textit{harmonic} if $\nabla\omega = 0$ and both \textit{exact} $d\omega=0$ and \textit{coexact} $\delta\omega=0$, such that $\mc H^p M = \setc{\nabla\omega = 0}{\omega\in \Omega^pM}$.
	Hodge \cite{ivancevic}:
	$\Omega^pM=\mc H^pM\oplus\im{d\Omega^{p-1}M}\oplus\im{\partial\Omega^{p+1}M}$.
	Note: $\nabla\omega=-\omega\nabla, \nabla^2\omega=\omega\nabla^2$ for higher-order tensor fields!
\end{theorem}

@@

\begin{lstlisting}[language = Julia]
using Grassmann, Makie
basis"2" # Euclidean
streamplot(vectorfield(exp(π*v12/2)),-1.5..1.5,-1.5..1.5)
streamplot(vectorfield(exp((π/2)*v12/2)),-1.5..1.5,-1.5..1.5)
streamplot(vectorfield(exp((π/4)*v12/2)),-1.5..1.5,-1.5..1.5)
streamplot(vectorfield(v1*exp((π/4)*v12/2)),-1.5..1.5,-1.5..1.5)
@basis S"+-" # Hyperbolic
streamplot(vectorfield(exp((π/8)*v12/2)),-1.5..1.5,-1.5..1.5)
streamplot(vectorfield(v1*exp((π/4)*v12/2)),-1.5..1.5,-1.5..1.5)
\end{lstlisting}
\begin{figure}[ht]
	\centering
	%\caption{Euclidean vector fields based on exponential operator:}
	\begin{subfigure}[b]{0.23\textwidth}
		\includegraphics[width=\textwidth]{img/plane-1.png}
		\caption{$x\oslash e^{\pi v_{12}/2}$, $\mb R^2$}
		%\label{fig:sample_figure}
	\end{subfigure}
	~
	\begin{subfigure}[b]{0.23\textwidth}
		\includegraphics[width=\textwidth]{img/plane-2.png}
		\caption{$ x\oslash e^{\frac\pi2 v_{12}/2}$, $\mb R^2$}
		%\label{fig:sample_figure}
	\end{subfigure}

@@

As a result of Grassmann's exterior \& interior products, the Hodge-DeRahm chain complex from cohomology theory is
\begin{align*}
	0 &\os d{\us\partial\rightleftarrows} \Omega^0(M) \os d{\us\partial\rightleftarrows} \Omega^1(M) \os d{\us\partial\leftrightarrows} \cdots \os d{\us\partial\rightleftarrows} \Omega^n(M) \os d{\us\partial\rightleftarrows} 0,
\end{align*}
having dimensional equivalence brought by the Grassmann-Poincare-Hodge complement duality,
\begin{align*}
	\mc H^{n-p}M &\cong \frac{\ker(d\Omega^{n-p}M)}{\im{d\Omega^{n-p+1}M}}, & \dim\mc H^pM &= \dim\frac{\ker(\partial\Omega^pM)}{\im{\partial\Omega^{p+1}M}}
\end{align*}
The rank of the grade $p$ boundary incidence operator is
$$ \text{rank}\gen{\partial\gen{M}_{p+1}}_p = \min\set{\dim\gen{\partial\gen{M}_{p+1}}_p,\dim\gen{M}_{p+1}} $$
Invariant topological information can be computed using the rank of homology groups, where $b_p(M)=\dim\mc H^pM$ are
$$ b_p(M) = \dim\gen{M}_{p+1} - \text{rank}\gen{\partial\gen{M}_{p+1}}_p - \text{rank}\gen{\partial\gen{M}_{p+2}}_{p+1} $$
Betti numbers with Euler characteristic $\chi(M) = \sum_p (-1)^pb_p$.

@@

In Fig. 4, different possible discrete bivector topologies in a projective Riemann sphere setting are examined. 
The figures are based on the product topology of two rotation bivectors.
When the Euclidean $\mb R^4$ basis is combined with projective geometric algebra, resulting one parameter Lie groups can be visualized as a fibration of a torus in $\mb R^3$. 
When the fourth $v_\infty$ basis direction is rotated into the Minkowski plane, the double rotation becomes a helix with translational single rotation. In examples (b)$\sim$(c) the bivector is modulated.
